# Volcano plots

```{r, echo = FALSE}
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```

The **Volcano plot** is a useful way to display the fold change and significance of each gene in the comparison. The **x-axis** is the Log2 Fold change, which shows the increase and decrease of genes within the current comparison. The values that are deviated from zero mean larger expression changes. The **y-axis** is the -Log10(Padj.Value), and the higher value corresponds to a smaller adjusted P-value, i.e., higher significace.

This volcano plot reveals the differentially expressed genes after running the DE analysis by EArun. By default, the Fold change Cutoff is set to 1.2, and the adjusted P-value Cutoff is set to 0.01.

The **`r colorize("Red dots", "red")`** are indicative of all genes that pass the Fold Change and adjusted P-value cutoff, while the **`r colorize("Green dots", "green")`** are indicative of genes that only pass the adjusted P-value cutoff. The **`r colorize("Grey dots", "grey")`** are usually in the middle, and they are not significant genes. By default, fifty random genes are labeled for both the up-regulated and down-regulated genes.

Moreover, this chapter will display volcano plots of all comparison groups.

```{r, echo=FALSE, fig.width = 12, fig.asp = 0.9, results = "asis"}
#Retrieve information from the previously saved bookdown.info.txt file
suppressMessages(library(knitr))
suppressMessages(library(kableExtra))

PWD <- system("pwd", intern = TRUE)
Info <- paste0(PWD, "/bookdown.info.txt")
Data <- read.table(Info)
Source_code <- Data$V2[1]
Working_dir <- Data$V2[2]
Config_file <- Data$V2[3]

#load Plot_Functions.R
setwd(Source_code)   #This is critical!
PlotFunctions <- paste0(Source_code, "Plot_Functions.R")
suppressWarnings(source("Plot_Functions.R"))

#Extract information from the Config_file
config <- sapply(yaml::read_yaml(Config_file),unlist)
sysConfig <- yaml::read_yaml(paste0(Source_code,"sys.yml"))
ProjectName <- config$prj_name

#Parsing the data
load(paste0(Working_dir,"../",ProjectName,".RData"))
ComparisonGroups <- as.character(unique(results_long$test))
if (length(ComparisonGroups) == 1){
  cat("There is only ", length(ComparisonGroups), " comparison in this project.\n\n")
}else{
  cat("There are total ", length(ComparisonGroups), " comparisons in this project.\n\n")
}

kable(comp_info, caption = "Comparisons", align = "c", booktabs = TRUE) %>% 
  kable_styling() %>%
  kableExtra::scroll_box(width = "100%")

#Loop through all comparisons
for (i in 1:length(ComparisonGroups)){
   cat('\n')  
   cat("## Comparison: ", ComparisonGroups[i], "\n") 
   p <- Volcano_Plot(results_long, ProteinGeneName, comparison = ComparisonGroups[i])
   print(p)
   cat('\n') 
}
```
