```{r, echo=FALSE}

#Retrieve information from the previously saved bookdown.info.txt file
PWD <- system("pwd", intern = TRUE)
Info <- paste0(PWD, "/bookdown.info.txt")
Data <- read.table(Info)
Source_code <- Data$V2[1]
Working_dir <- Data$V2[2]
Config_file <- Data$V2[3]

#Extract information from the Config_file
config <- sapply(yaml::read_yaml(Config_file),unlist)
sysConfig <- yaml::read_yaml(paste0(Source_code,"sys.yml"))
ProjectName <- config$prj_name

load(paste0(Working_dir,"../",ProjectName,".RData"))
NumberOfSamples <- dim(data_wide)[2]
NumberOfGroups <- length(unique(MetaData$group))
NumberOfGenes <- dim(data_wide)[1]
NumberOfComparisons <- dim(comp_info)[1]
```

--- 
title: Bookdown report of `r ProjectName`
date: "Generated on `r Sys.Date()`"
site: bookdown::bookdown_site
---

# Introduction {-}

This is a summary report generated by EAreport (part of ExpressionAnalysis) using Bookdown.

This report contains basic plots and tables generated by the EA pipeline.

### Project overview {-}

Here is a brief overview of the project:

- Project name: **`r config$prj_name`**

- Species: **`r config$species`**

- Description: **`r config$shinyOne_Description`**

- Number of samples: **`r NumberOfSamples`**

- Number of groups: **`r NumberOfGroups`**

- Number of Genes/Proteins: **`r NumberOfGenes`**

- Number of comparisons: **`r NumberOfComparisons`**

- Number of Samples in Each Group:

```{r, echo=FALSE}
suppressMessages(library(knitr))
suppressMessages(library(kableExtra))
Counting <- table(MetaData$group)
Table <- as.data.frame(matrix(nrow = 1,data = as.vector(Counting)))
colnames(Table) <- names(Counting)
row.names(Table) <- "count"

kable(Table, align = "c", booktabs = TRUE) %>%
  kable_styling() %>%
  kableExtra::scroll_box(width = "100%")
```

### Project files {-}

The ExpressionAnalysis (EA) pipeline analyzes RNAseq data provided by the user, or from GEO database, and generates results for interactive visualization. This report is part of ExpressionAnalysis (EA) output, and contains key results of the project.

For more flexible and interactive visualization, please upload the following Rdata files to [Quickomics](http://quickomics.bxgenomics.com/).

**`r ProjectName`_estCount.rds**

**`r ProjectName`_network.RData**

**`r ProjectName`.RData**

**`r ProjectName`.csv**

The following sections are included in this bookdown report:

1, QC plots

2, Covariates

3, PCA plot

4, Volcano plots

5, Heatmap plots
