---
title: "EA Slide Deck"
output: 
  revealjs::revealjs_presentation:
    self_contained: false
    transition: none
    reveal_plugins: ["search", "zoom", "notes", "chalkboard", "menu"]
    reveal_options:
      chalkboard: 
        theme: chalkboard
        toggleNotesButton: true
      menu:
        numbers: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=5, dev='svg', echo=FALSE, warning=FALSE, message=FALSE)
```

```{css css-settings, echo=FALSE}
.reveal section img {
  border: 0px;
}

```

```{r, echo=FALSE}
#Retrieve information from the previously saved bookdown.info.txt file
info_txt <- "/home/whu1/EA_results/Ea_projects/RNASequest/example/SRP199678/EA20220408_0/BookdownReport/bookdown.info.txt"
Data <- read.table(info_txt)
Source_code <- Data$V2[1]
Working_dir <- Data$V2[2]
Config_file <- Data$V2[3]

#A copy of the useAlias function, without printing out the D
useAliasCopy <- function(config,D){
    if(!is.null(config$sample_alias)){
        message("Applying alias")
        colnames(D$counts) <- rownames(D$meta) <- D$meta[,config$sample_alias]
        if(!is.null(D$effLength)) colnames(D$effLength) <- rownames(D$meta)
        if(!is.null(D$logTPM)) colnames(D$logTPM) <- rownames(D$meta)
        if(!is.null(D$seqQC)) rownames(D$seqQC) <- rownames(D$meta)
    }
    return(D)
}


#Extract information from the Config_file
config <- sapply(yaml::read_yaml(Config_file),unlist)
sysConfig <- yaml::read_yaml(paste0(Source_code,"sys.yml"))
ProjectName <- config$prj_name
load(paste0(Working_dir,"../",ProjectName,".RData"))
NumberOfSamples <- dim(data_wide)[2]
NumberOfGroups <- length(unique(MetaData$group))
NumberOfGenes <- dim(data_wide)[1]
NumberOfComparisons <- dim(comp_info)[1]
EAOutputDir <- substr(Working_dir, 1, nchar(Working_dir)-15)

# load Plot_Functions.R
setwd(Source_code)
PlotFunctions <- paste0(Source_code, "Plot_Functions.R")
source("utility.R")
source("Plot_Functions.R")

suppressMessages(checkConfig(config))
D <- suppressMessages(getEAData(config))
D <- useAliasCopy(config,D)
```


## Navigation buttons
<ul>
<li> Press "Esc" or "o" key to enter the the slide overview
<li> Alt + mouse click to zoom in/out
<li> Press "s" key to speaker view to see notes
</ul>
```{r echo=FALSE}
htmltools::tags$aside('Here are some notes.',class = "notes")
```


## Quality control (QC) plots
QC: gene ratios. Top 1% gene of TPM

```{r, echo=FALSE}
## Setting plotting parameters ----
estT <- 2^D$logTPM-config$count_prior
estC=D$counts
qc=D$seqQC
prioQC=sysConfig$qc2meta
gInfo=D$gInfo
replot=config$seqQC_replot
topN=c(1,10,30)
# Generating ratio plots
if(!is.null(gInfo) & sum(rownames(gInfo)!=gInfo$Gene.Name,na.rm=T)>0){
  rownames(estT) <- paste(rownames(estT),gInfo[rownames(estT),"Gene.Name"],sep="|")
  if(!is.null(estC)) rownames(estC) <- paste(rownames(estC),gInfo[rownames(estC),"Gene.Name"],sep="|")
}
pdfW <- max(ncol(estT)/10+2,6)
topN <- setNames(topN,paste0("Top",topN))
D <- t(apply(estT,2,function(x){
  x <- sort(x,decreasing=T)
  return(sapply(topN,function(i)return(sum(x[1:i])/sum(x)*100)))
}))
D <- cbind(sID=factor(rownames(D),levels=rownames(D)),data.frame(D))
i <- colnames(D)[2]
p1 <- ggplot(D,aes_string(x="sID",y=i))+
          geom_bar(stat="identity")+
          xlab("")+ylab("% of Total TPM")+
          ggtitle(paste(i,"genes"))+
          theme_minimal()+
          theme(axis.text.x=element_text(angle=90,vjust=0.5,hjust=1),
                legend.position = "none")
p1
```

## Quality control (QC) plots
QC: gene ratios. Top 10% gene of TPM

```{r, echo=FALSE}
i <- colnames(D)[3]
p2 <- ggplot(D,aes_string(x="sID",y=i))+
          geom_bar(stat="identity")+
          xlab("")+ylab("% of Total TPM")+
          ggtitle(paste(i,"genes"))+
          theme_minimal()+
          theme(axis.text.x=element_text(angle=90,vjust=0.5,hjust=1),
                legend.position = "none")
p2
```

## Quality control (QC) plots
QC: gene ratios. Top 30% gene of TPM

```{r, echo=FALSE}
i <- colnames(D)[4]
p3 <- ggplot(D,aes_string(x="sID",y=i))+
          geom_bar(stat="identity")+
          xlab("")+ylab("% of Total TPM")+
          ggtitle(paste(i,"genes"))+
          theme_minimal()+
          theme(axis.text.x=element_text(angle=90,vjust=0.5,hjust=1),
                legend.position = "none")
p3
```

## Quality control (QC) plots
QC: top expressed genes. Top gene by TPM

```{r, echo=FALSE}
topUnion <- 10
p4 <- plotTopGeneRatio(estT, 10)
p4
```

## Quality control (QC) plots
QC: top expressed genes. Top gene by counts

```{r, unsmooth_rc}
p5 <- plotTopGeneRatio(estC,10,selG=levels(p4$data[,2]))+
    xlab("% of total counts")+ggtitle(p4$labels$title)
p5
```

## Covariates plots

```{r, echo=FALSE}
suppressMessages(library(knitr))
suppressMessages(library(kableExtra))
options(width = 100)
res <- suppressWarnings(suppressMessages(PC_covariates_out(data_wide, MetaData, N_col = 3)))
sig_covariates=res$selVar_All

i <- 1
pc <- sig_covariates$PC[i]
var <- sig_covariates$Covariate[i]
add_text=TRUE

data.all=res$data.all
selVar=res$selVar_All
  
Num_names=names(select_if(data.all, is.numeric))
  
if (var %in% Num_names) {
    p6<-suppressWarnings(ggplot(data.all, aes(x=!!sym(var), y=!!sym(pc)) )+geom_point()+
      stat_summary(fun.data= mean_cl_normal) + geom_smooth(formula = y ~ x, method='lm')+theme_half_open())
  } else {
    p6<-suppressWarnings(ggplot(data.all,  aes(x=!!sym(var), y=!!sym(pc)) )+geom_boxplot()+geom_jitter(alpha=0.7, width=0.1)+theme_half_open() +
      theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=0.5)))
  }
p6
```

## Covariates plots

```{r, echo=FALSE}
i <- 2
pc <- sig_covariates$PC[i]
var <- sig_covariates$Covariate[i]
add_text=TRUE
data.all=res$data.all
selVar=res$selVar_All
  
Num_names=names(select_if(data.all, is.numeric))
  
if (var %in% Num_names) {
    p7<-suppressWarnings(ggplot(data.all, aes(x=!!sym(var), y=!!sym(pc)) )+geom_point()+
      stat_summary(fun.data= mean_cl_normal) + geom_smooth(formula = y ~ x, method='lm')+theme_half_open())
  } else {
    p7<-suppressWarnings(ggplot(data.all,  aes(x=!!sym(var), y=!!sym(pc)) )+geom_boxplot()+geom_jitter(alpha=0.7, width=0.1)+theme_half_open() +
      theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=0.5)))
  }
p7
```

## PCA plots
Regular PCA plot

```{r, echo=FALSE}
par(mar = c(0.1, 0.1, 0.1, 0.1))
PCA_Plot(data_wide, MetaData)
```


## PCA plots
PCA of covariates

```{r, echo=FALSE}
res <- suppressWarnings(suppressMessages(PC_covariates_out(data_wide, MetaData, N_col=1)))
PCA_plots<-PCA_sig_covariates(res)
i = 1
plot_name <- names(PCA_plots[i])
plot <- suppressWarnings(PCA_plots[[i]])
suppressWarnings(print(ggdraw(plot)))

```

## PCA plots
PCA of covariates

```{r, echo=FALSE}
i = 2
plot_name <- names(PCA_plots[i])
plot <- suppressWarnings(PCA_plots[[i]])
suppressWarnings(print(ggdraw(plot)))
```

## Volcano plots
To visulize the fold change and significance of DEGs in a comparison.

```{r, echo=FALSE}
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}

ComparisonGroups <- as.character(unique(results_long$test))

#Use the results_long from RData to get DEG counts
FCcut=log2(1.2) #Fold change cutoff, 1.2 fold as default
pvalcut=0.05 #FDR cutoff  
tmpdat = results_long %>% dplyr::filter( Adj.P.Value < pvalcut & abs(logFC) > FCcut)
deg_stat<-tmpdat%>%group_by(test)%>%dplyr::summarize(DEG=n(), Up=sum(logFC>0), Down=sum(logFC<0))%>%ungroup
names(deg_stat)[1]="Comparison"

i <- 1
p <- Volcano_Plot(results_long, ProteinGeneName, comparison = ComparisonGroups[i])
print(p)

```

## Heatmap plots
Heatmap of the top 1000 most variable genes
```{r, echo=FALSE}
p <- Heatmap_Plot(data_wide, MetaData, results_long, ProteinGeneName, heatmap_subset = "variable", maxgenes = 1000)
draw(p, merge_legend = TRUE, auto_adjust = FALSE)
```

## Heatmap plots
Heatmap of group comparisons

```{r, echo=FALSE}
i <- 1
try(p <- Heatmap_Plot(data_wide, MetaData, results_long, ProteinGeneName, heatmap_subset = "subset", 
                 comparison = ComparisonGroups[i],
                 maxgenes = 1000))
try(draw(p, merge_legend = TRUE, auto_adjust = FALSE))

```
